---
title: "Primera entrega Lab ADEI"
author: "Alejandro Alarcón"
date: "10/9/2021"
output: pdf_document
---

# Set directory and load libraries to work with used car data:
```{r setup, include=FALSE, echo=FALSE}
require("knitr")
opts_knit$set(root.dir = "/Users/aleibz/ADEI/CarPrices/")
```

```{r}
setwd("/Users/aleibz/ADEI/CarPrices/")
#install.packages("car")
#install.packages("FactoMineR")
#install.packages("knitr")
library(car)
library(FactoMineR)
library(knitr)
```

# Load the Data and Select a random subset

We will proceed to load the data and unify it in a single dataframe. After that, we will select a subset of 5000 elements at random and save it in an .RData file.

```{r}
# Lecture of DataFrames:
df1 <- read.table("audi.csv",header=T, sep=",")
df1$manufacturer <- "Audi"
df2 <- read.table("bmw.csv",header=T, sep=",")
df2$manufacturer <- "BMW"
df3 <- read.table("merc.csv",header=T, sep=",")
df3$manufacturer <- "Mercedes"
df4 <- read.table("vw.csv",header=T, sep=",")
df4$manufacturer <- "VW"

# Union by row:
df <- rbind(df1,df2,df3,df4)
dim(df)  # Size of data.frame
str(df) # Object class and description
names(df) # List of variable names

### Use birthday of 1 member of the group as random seed:
set.seed(20071998)
# Random selection of x registers:
sam<-as.vector(sort(sample(1:nrow(df),5000)))
head(df)  # Take a look to the first rows/instances (6 rows)
df<-df[sam,]  # Subset of rows _ It will be my sample
summary(df)

#Keep information in an .Rdata file:
save(list=c("df"),file="MyOldCars-Raw.RData")

```

## Load Required Packages

Now that we have loaded the data and generated a 5000 elements dataset we will proceed to load the packages for further exploration of this dataset.

```{r}
options(contrasts=c("contr.treatment","contr.treatment"))

requiredPackages <- c("effects","FactoMineR","car", "factoextra","RColorBrewer","ggplot2","dplyr","ggmap","ggthemes","knitr")

package.check <- lapply(requiredPackages, FUN = function(x) {
  if (!require(x, character.only = TRUE)) {
    install.packages(x, dependencies = TRUE)
    library(x, character.only = TRUE)
  }
})
#verify they are loaded
search()
```

```{r}
# Clear plots
if(!is.null(dev.list())) dev.off()

# Clean workspace
rm(list=ls())

setwd("/Users/aleibz/ADEI/CarPrices/")
filepath<-"/Users/aleibz/ADEI/CarPrices/"

load(paste0(filepath,"MyOldCars-Raw.RData"))
```

## Useful Functions

We will be using some predefined functions to make the code more readable and easy.

```{r, echo=FALSE}

# Some useful functions
calcQ <- function(x) {
  s.x <- summary(x)
  iqr<-s.x[5]-s.x[2]
  list(souti=s.x[2]-3*iqr, mouti=s.x[2]-1.5*iqr, min=s.x[1], q1=s.x[2], q2=s.x[3], 
       q3=s.x[5], max=s.x[6], mouts=s.x[5]+1.5*iqr, souts=s.x[5]+3*iqr ) }

countNA <- function(x) {
  mis_x <- NULL
  for (j in 1:ncol(x)) {mis_x[j] <- sum(is.na(x[,j])) }
  mis_x <- as.data.frame(mis_x)
  rownames(mis_x) <- names(x)
  mis_i <- rep(0,nrow(x))
  for (j in 1:ncol(x)) {mis_i <- mis_i + as.numeric(is.na(x[,j])) }
  list(mis_col=mis_x,mis_ind=mis_i) }

countX <- function(x,X) {
  n_x <- NULL
  for (j in 1:ncol(x)) {n_x[j] <- sum(x[,j]==X) }
  n_x <- as.data.frame(n_x)
  rownames(n_x) <- names(x)
  nx_i <- rep(0,nrow(x))
  for (j in 1:ncol(x)) {nx_i <- nx_i + as.numeric(x[,j]==X) }
  list(nx_col=n_x,nx_ind=nx_i) }

```

## Factors

We will now proceed to generate factors from categorical variables.

```{r}
summary(df)
names(df)

#Model
df$model <- factor(paste0(df$manufacturer, "-", df$model))
levels(df$model)

#Transmission
df$transmission <- factor( df$transmission )
levels( df$transmission )
df$transmission <- factor( df$transmission, levels = c("Manual","Semi-Auto","Automatic"),labels = paste0("f.Trans-",c("Manual","SemiAuto","Automatic")))
head( df )

#FuelType
df$fuelType <- factor(df$fuelType)
levels(df$fuelType)
df$fuelType <- factor( df$fuelType, levels = c("Diesel","Petrol","Hybrid"), labels = paste0("f.Fuel-",c("Diesel","Petrol","Hybrid")))

#Manufacturer
df$manufacturer <- factor(df$manufacturer)

# Falta pasar year a factor y generar la columna Age como 2021 - year
    #Probablemente tendriamos que hacer esto despues de hacer todo lo que es la imputacion y los missings y todo

# Puede que también haya que pasar la variable EngineSize a factor, ya que no es una variable continua


# Remove uninformative variables from data.frame if present
```


# Initialization of counts for missings, outliers and errors. All numerical variables have to be checked before

```{r}

#######################################################
imis<-rep(0,nrow(df))  # rows - trips
jmis<-rep(0,2*ncol(df))  # columns - variables
######################################################
mis1<-countNA(df)
imis<-mis1$mis_ind
imis
mis1$mis_col # Number of missings for the current set of variables

#######################################################
iouts<-rep(0,nrow(df))  # rows - trips
jouts<-rep(0,2*ncol(df))  # columns - variables
######################################################

#######################################################
ierrs<-rep(0,nrow(df))  # rows - trips
jerrs<-rep(0,2*ncol(df))  # columns - variables
######################################################
```

As we can observe in the mis1 variable, there appear to be missings in just one of the colums of our dataframe.

# Mostrar el nombre de la columna y las filas enteras para las cuales aparecen los valores missing

## Data Coding and Clearance

Create new variables derived from the original ones.

```{r}
summary(df$model)
# It is a categorical(=factor) variable   NO PROBLEM

####   Variable df$year
summary(df$year)
df$age <-  2021 - df$year 
df$year<-factor(df$year)
# It is a categorical(=factor) variable   NO PROBLEM

#### Variable df$age
summary(df$age)

####   Variable df$price
summary(df$price)

####   Variable df$transmission
summary(df$transmission)

####  Variable mileage
summary(df$mileage)

#### Take a look at each variable
###   Variable df$engineSize
# 0.00 seems to be an error for non-electric vehicles
# Peak at individuals with this "0" value:
df[which(df[,"engineSize"]==0),]

# It is a quantitive variable  Non-possible values will be recoded to NA
sel<-which(df$engineSize ==0)
ierrs[sel]<-ierrs[sel]+1
sel                 #### sel contains the rownames of the individuals with "0" 
#                        as  value for engineSize
# df[sel,"engineSize"]<-3    # non-possible values are replaced by NA, missing value symbol in R
df[sel,"engineSize"]<-NA
selmiss <- sel
##### 

# Outlier detection
Boxplot(df$price)
var_out<-calcQ(df$price)

#Estas dos líneas no funcionan
  #abline(h=var_out$souts,col="red")
  #abline(h=var_out$souti,col="red")

# Outliers:
llout<-which((df$price<0)|(df$price>60000)) #Estos valore estan hardcoded y molaría pasarlos como parámetro
iouts[llout]<-iouts[llout]+1
# Assignation of an NA to my response variable:
df[1,"price"]<-NA 

# Do that for all the variables....

```


